Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-26 article.tex
modified tab 1 + correction

\begin{table}[t]
\caption{Evaluation of different methods on the four \textit{normal} datasets\label{tab:comparison}}
\centering
\small
\begin{tabular}{@{}l *{8}{c}@{}}
\hline
  & \multicolumn{4}{c}{AUROC} & \multicolumn{4}{c}{AUPRC} \\
\hline
Methods $\backslash$ normal- & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 \\
No  filtering       & 0.777 & 0.767 & 0.772 & 0.774 & 0.070 & 0.064 & 0.068 & 0.072\\
$ h \circ g \circ f_1$                 & 0.925 & 0.925 & 0.924 & 0.923 & 0.322 & 0.324 & 0.323 & 0.312\\
$ w \circ h \circ g \circ f_1$               & 0.931 & 0.930 & 0.928 & 0.927 & 0.333 & 0.334 & 0.327 & 0.313\\
+ PCA         & 0.932 & 0.931 & 0.930 & 0.928 & 0.364 & 0.366 & 0.359 & 0.344\\
Averaging           & 0.943 & 0.942 & 0.942 & 0.939 & 0.403 & 0.404 & 0.398 & 0.388\\
Full method &\\
\hline
%Averaging + Directivity & 0.944 & 0.943 & 0.942 & 0.940 & 0.404 & 0.405 & 0.399 & 0.389\\
Correlation &\\
GTE & \\
GENIE3 &\\
\end{tabular}
\caption{TODO \textcolor{red}{version simple!!}}
\label{tab:tab3}
\end{table}

@@

In the bottom part of Table \ref{tab:comparison}, the proposed
approach is compared with three other methods: standard (Pearson)
correlation (PC), generalized transfer entropy (GTE), and GENIE3. ROC
and PR curves on the \textit{normal-2} network are shown for all
methods in Figure \ref{fig:}. Pearson correlation measures the
unconditional linear (in)dependence between variables and it should
thus not be able to filter out indirect interactions between
neurons. GTE \citep{stetter2012model} was proposed as a baseline for
the challenge. This method builds on Transfer Entropy to measure the
association between two neurons. Unlike our approach, it can predict
the direction of the edges. GENIE3 \citep{huynhthu2010inferring} is a
gene regulatory network inference method that was the best performer
of the DREAM5 challenge \cite{dream5paper}. When transposed to neural
networks, this method uses as a confidence score for the edge going
from neuron $i$ to neuron $j$ the importance score of variable $X_i$
in a Random Forests model trying to predict $X_j$ from all variables
in $X\setminus X_j$. To reduce computing times of this method, we had
to reduce each tree in the Random Forests model to a maximum depth of
3. This constraint potentially severly affects the performance of this
method with respect to the use of fully grown trees. We nevertheless
provide these results for comparison purpose. PC and GENIE3 were both
applied on the same pre-filtered time series as partial
correlation. For GENIE3, we built 10000 trees per neuron and we used
default settings for all other parameters (except for the maximal tree
depth). For GTE, we reproduced the exact same setting (conditioning
level and pre-processing) that was used for the challenge.

@@

\begin{align}
s_{ij} = \sum_{t=1}^{T - 1} \mathbb{1}(x_j^{t+1} - x_i^t \in \left[\phi_1, \phi_2\right])
\end{align}
where $\phi_1$ and $\phi_2$ are parameters of the heuristic.

@@

\begin{align}
z_{ij} = \sum_{t=1}^{T - 1}
    \mathbb{1}(x_j^{t+1} - x_i^t \in \left[\phi_1, \phi_2 \right]) -
    \mathbb{1}(x_i^{t+1} - x_j^t \in \left[\phi_1, \phi_2 \right]),
\end{align}
where $T$ is the number of time steps, $\phi_1$ and $\phi_2$ are parameters.

@@

% Introduce a custom solution to make the Y matrix asymmetric
% Introduce directivity
\textcolor{red}{Est-ce qu'on ne le ferait pas passer dans l'appendice B? Ou du moins l'équation?}
Since partial correlation is a symmetric measure, the causal mechanism behind the
interaction ($A$ directly causes $B$, $B$ directly causes $A$ or both) can not
be directly inferred\footnote{Also note that two other causal mechanisms might be
implied while having a non-zero partial coefficient: (i) a pair of variables
is induced by a common (hidden) variable; (ii) $A$ (respectively $B$) is
conditionally correlated to a hidden variable affecting $B$ (resp. $A$)
\citep{de2004discovery}.}.
In order to retrieve causal information, we developed an
heuristics based on the following observation. The rise of fluorescence
indicates the activation of a neuron $i$. If a neuron $j$ have also
an increased of fluorescence after a slight delay, this could have been
caused by the neuron $i$. Thus, we count whenever
such events occurred between consecutive time steps
\[
\hat{Z}: z_{ij} = \sum_{t=1}^{T - 1}
    \mathbb{1}(x_j^{t+1} - x_i^t \in \left[f_1, f_2 \right]) -
    \mathbb{1}(x_i^{t+1} - x_j^t \in \left[f_1, f_2 \right]), \forall i, j \in V
\]
where $f_1$ and $f_2$ are parameters of the method.
Finally, the weighted average of correlation matrices is combined with $Z$ such
that $Z$ has a contribution of $0.3\%$ to the prediction of an edge.

>>>>

\begin{table}[t]
\caption{Evaluation of different methods on the four \textit{normal} datasets\label{tab:comparison}}
\centering
\small
\begin{tabular}{@{}l *{8}{c}@{}}
\hline
  & \multicolumn{4}{c}{AUROC} & \multicolumn{4}{c}{AUPRC} \\
\hline
Methods $\backslash$ normal- & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 \\
No  filtering       					& 0.777 & 0.767 & 0.772 & 0.774 & 0.070 & 0.064 & 0.068 & 0.072\\
$ h \circ g \circ f_1$                  & 0.923 & 0.925 & 0.923 & 0.922 & 0.311 & 0.315 & 0.313 & 0.304\\
$ w \circ h \circ g \circ f_1$          & 0.931 & 0.929 & 0.928 & 0.926 & 0.326 & 0.323 & 0.319 & 0.303\\
+ PCA         							& 0.932 & 0.930 & 0.928 & 0.926 & 0.355 & 0.353 & 0.350 & 0.333\\
Averaging           					& 0.938 & 0.936 & 0.936 & 0.932 & 0.391 & 0.389 & 0.385 & 0.374\\
Full method           					& 0.943 & 0.942 & 0.942 & 0.939 & 0.403 & 0.404 & 0.398 & 0.388\\
\hline
Correlation & 0.886 & 0.884 & 0.891 &  0.877 & 0.153 & 0.145 & 0.170 & 0.132\\
GTE & 0.890 & 0.893 & 0.894 & 0.873 & 0.171 & 0.174 & 0.197 & 0.142\\
GENIE3 &\\
\end{tabular}
\caption{TODO \textcolor{red}{version simple!!}}
\label{tab:tab3}
\end{table}

@@

In the bottom part of Table \ref{tab:comparison}, the proposed
approach is compared with three other methods: standard (Pearson)
correlation (PC), generalized transfer entropy (GTE), and GENIE3. ROC
and PR curves on the \textit{normal-2} network are shown for all
methods in Figure \ref{fig:}. Pearson correlation measures the
unconditional linear (in)dependence between variables and it should
thus not be able to filter out indirect interactions between
neurons. GTE \citep{stetter2012model} was proposed as a baseline for
the challenge. This method builds on Transfer Entropy to measure the
association between two neurons. Unlike our approach, it can predict
the direction of the edges. GENIE3 \citep{huynhthu2010inferring} is a
gene regulatory network inference method that was the best performer
of the DREAM5 challenge \cite{dream5paper}. When transposed to neural
networks, this method uses as a confidence score for the edge going
from neuron $i$ to neuron $j$ the importance score of variable $X_i$
in a Random Forests model trying to predict $X_j$ from all variables
in $X\setminus X_j$. To reduce computing times of this method, we had
to reduce each tree in the Random Forests model to a maximum depth of
3. This constraint potentially severly affects the performance of this
method with respect to the use of fully grown trees. We nevertheless
provide these results for comparison purpose. PC and GENIE3 were both
applied on the same pre-filtered time series ($g \circ f_1$). For GENIE3, we built 10000 trees per neuron and we used
default settings for all other parameters (except for the maximal tree
depth). For GTE, we reproduced the exact same setting (conditioning
level and pre-processing) that was used for the challenge.

@@

\begin{align}
s_{ij} = \sum_{t=1}^{T - 1} \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2\right])
\end{align}
where $\phi_1$ and $\phi_2$ are parameters of the heuristic.

@@

\begin{align}
z_{ij} = \sum_{t=1}^{T - 1}
    \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2 \right]) -
    \mathbb{1}((x_i^{t+1} - x_j^t) \in \left[\phi_1, \phi_2 \right]),
\end{align}
where $T$ is the number of time steps, $\phi_1$ and $\phi_2$ are parameters.

@@

% Introduce a custom solution to make the Y matrix asymmetric
% Introduce directivity
\textcolor{red}{Est-ce qu'on ne le ferait pas passer dans l'appendice B? Ou du moins l'équation?}
Since partial correlation is a symmetric measure, the causal mechanism behind the
interaction ($A$ directly causes $B$, $B$ directly causes $A$ or both) can not
be directly inferred\footnote{Also note that two other causal mechanisms might be
implied while having a non-zero partial coefficient: (i) a pair of variables
is induced by a common (hidden) variable; (ii) $A$ (respectively $B$) is
conditionally correlated to a hidden variable affecting $B$ (resp. $A$)
\citep{de2004discovery}.}.
In order to retrieve causal information, we developed an
heuristics based on the following observation. The rise of fluorescence
indicates the activation of a neuron $i$. If a neuron $j$ have also
an increased of fluorescence after a slight delay, this could have been
caused by the neuron $i$. Thus, we count whenever
such events occurred between consecutive time steps
\[
\hat{Z}: z_{ij} = \sum_{t=1}^{T - 1}
    \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[f_1, f_2 \right]) -
    \mathbb{1}((x_i^{t+1} - x_j^t) \in \left[f_1, f_2 \right]), \forall i, j \in V
\]
where $f_1$ and $f_2$ are parameters of the method.
Finally, the weighted average of correlation matrices is combined with $Z$ such
that $Z$ has a contribution of $0.3\%$ to the prediction of an edge.
